---
title: "Growth Model Exploration"
author: "Alice Stears"
date: "3/10/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Load packages
require(effects)
require(lattice)
require(latticeExtra)
require(gridExtra)
require(grid)
require(ggpubr)
require(lme4)
require(tidyverse)
require(cowplot)
require(lmerTest)
require(stargazer)
require(ggeffects)
require(sjPlot)
```

```{r, echo = FALSE}
#set the path to the name of the file containing the 'scripts' folder
path <-"/Users/Alice/Dropbox/Grad School/Research/Trait Project/CO_sgs Analysis/CO-Sgs-paper"
setwd(path)
#get model result data into the environment
load("./scripts/script4_output.RData") #change the file name to the most current version of model runs
```
# Growth Models using size$_{t+1}$ as the response variable
The following growth models use size$_{t+1}$ as the response variable, with both a random slope and a fixed effect of size$_{t}$. The figure below shows the updated growth model results for five traits. There is a significant interaction (albeit weak) between SPEI and LDMC for models of growth, but no other significant interactions. Including traits in all growth models actually decreases the AIC of the model in comparison to models without trait data.    

## Here is a figure showing survival and growth model results for both forbs and graminoids
```{r, echo = FALSE, fig.width=9, fig.height=10}
## make figure for graminoid survival

#get 2.5 and 97.5 percentiles of the distribution
meanSPEI_G <- mean(CO_grams$SPEI_s, na.rm = TRUE)
sdSPEI_G <- sd(CO_grams$SPEI_s, na.rm = TRUE)
#get 97.5 quantile of the distribution
SPEI_97_5_G <- qnorm(.975, meanSPEI_G, sdSPEI_G) 
SPEI_2_5_G <- qnorm(.025, meanSPEI_G, sdSPEI_G)

spei_vals <- c(SPEI_2_5_G, SPEI_97_5_G)

#for TLP_s
TLP_vals <- seq(min(CO_grams$TLP_s, na.rm = TRUE), max(CO_grams$TLP_s, na.rm = TRUE), length.out = 20)
TLP_G_dat<- ggpredict(mSurvTLP_grams, terms = c("TLP_s[TLP_vals]", "SPEI_s[spei_vals]"), type = "fixed", back.transform = TRUE, typical = "mean")
  #the ggpredict() function uses mean values for each of the fixed effects that aren't specified, and population-level values for the random effects

#for LDMC_s
LDMC_vals <- seq(min(CO_grams$LDMC_s, na.rm = TRUE), max(CO_grams$LDMC_s, na.rm = TRUE), length.out = 20)
LDMC_G_dat <- ggpredict(mSurvLDMC_grams, terms = c("LDMC_s[LDMC_vals]", "SPEI_s[spei_vals]"), type = "fixed", back.transform = TRUE)

#for RDMC_s
RDMC_vals <- seq(min(CO_grams$RDMC_s, na.rm = TRUE), max(CO_grams$RDMC_s, na.rm = TRUE), length.out = 20)
RDMC_G_dat <- ggpredict(mSurvRDMC_grams, terms = c("RDMC_s[RDMC_vals]", "SPEI_s[spei_vals]"), type = "fixed", back.transform = TRUE)

# for RDiam
RDiam_vals <- seq(min(CO_grams$RDiam_s, na.rm = TRUE), max(CO_grams$RDiam_s, na.rm = TRUE), length.out = 20)
RDiam_G_dat <- ggpredict(mSurvRDiam_grams, terms = c("RDiam_s[RDiam_vals]", "SPEI_s[spei_vals]"), type = "fixed", back.transform = TRUE)

# for SLA
SLA_vals <- seq(min(CO_grams$SLA_s, na.rm = TRUE), max(CO_grams$SLA_s, na.rm = TRUE), length.out = 20)
SLA_G_dat <- ggpredict(mSurvSLA_grams, terms = c("SLA_s[SLA_vals]", "SPEI_s[spei_vals]"), type = "fixed", back.transform = TRUE)

#make a data.frame to contain all of the values for each trait
GramDat <- data.frame(trait = c("scaled(Turgor Loss Point) (MPa)"), x = TLP_G_dat$x, GramSurv = TLP_G_dat$predicted, CI_low = TLP_G_dat$conf.low, CI_high = TLP_G_dat$conf.high, SPEI = TLP_G_dat$group, lab = "A")

GramDat <- rbind(GramDat, data.frame(trait = c("scaled(Leaf Dry Matter Content) (g/g)"), x = LDMC_G_dat$x, GramSurv = LDMC_G_dat$predicted, CI_low = LDMC_G_dat$conf.low, CI_high = LDMC_G_dat$conf.high, SPEI = LDMC_G_dat$group, lab = "B"))

GramDat <- rbind(GramDat, data.frame(trait = c("scaled(Root Dry Matter Content) (g/g)"), x = RDMC_G_dat$x, GramSurv = RDMC_G_dat$predicted, CI_low = RDMC_G_dat$conf.low, CI_high = RDMC_G_dat$conf.high, SPEI = RDMC_G_dat$group, lab = "C"))

GramDat <- rbind(GramDat, data.frame(trait = c("scaled(Average Root Diameter) (cm)"), x = RDiam_G_dat$x, GramSurv = RDiam_G_dat$predicted, CI_low = RDiam_G_dat$conf.low, CI_high = RDiam_G_dat$conf.high, SPEI = RDiam_G_dat$group, lab = "C"))

GramDat <- rbind(GramDat, data.frame(trait = c("scaled(Specific Leaf Area) (g/cm2)"), x = SLA_G_dat$x, GramSurv = SLA_G_dat$predicted, CI_low = SLA_G_dat$conf.low, CI_high = SLA_G_dat$conf.high, SPEI = SLA_G_dat$group, lab = "C"))


#levels(GramDat$trait) <- c("scaled(Turgor~Loss~Point)~(MPa)", "scaled(Leaf~Dry~Matter~Content)~(g/g)", "scaled(Root~Dry~Matter~Content)~(g/g)", "scaled(Average~Root~Diameter)~(cm)", expression('scaled(Specific~Leaf~Area)~(g/c'*m^'2'*')'))

#make data for rug plot
RugDat_G <-  data.frame(rug = CO_grams$TLP_s, trait = "scaled(Turgor Loss Point) (MPa)")
RugDat_G <- rbind(RugDat_G, data.frame(rug = CO_grams$LDMC_s, trait = "scaled(Leaf Dry Matter Content) (g/g)"))
RugDat_G <- rbind(RugDat_G, data.frame(rug = CO_grams$RDMC_s, trait = "scaled(Root Dry Matter Content) (g/g)"))
RugDat_G <- rbind(RugDat_G, data.frame(rug = CO_grams$RDiam_s, trait = "scaled(Average Root Diameter) (cm)"))
RugDat_G <- rbind(RugDat_G, data.frame(rug = CO_grams$SLA_s, trait = "scaled(Specific Leaf Area) (g/cm2)"))
#levels(RugDat_G$trait) <-c("scaled(Turgor~Loss~Point)~(MPa)", "scaled(Leaf~Dry~Matter~Content)~(g/g)", "scaled(Root~Dry~Matter~Content)~(g/g)", "scaled(Average~Root~Diameter)~(cm)", expression('scaled(Specific~Leaf~Area)~(g/c'*m^'2'*')'))

#text for labels
dat_text <- data.frame(
  label = c("A", "D", "G", "J", "M"),
  sig = c("*,§","*,§","*,§","*,§","*,§"),
  trait = c("scaled(Turgor Loss Point) (MPa)", "scaled(Leaf Dry Matter Content) (g/g)", "scaled(Root Dry Matter Content) (g/g)","scaled(Average Root Diameter) (cm)", "scaled(Specific Leaf Area) (g/cm2)"),
  x    = c(min(TLP_G_dat$x),min(LDMC_G_dat$x),min(RDMC_G_dat$x), min(RDiam_G_dat$x), min(SLA_G_dat$x)),
  y     = c(1,1,1,1,1),
  x1 = c(max(TLP_G_dat$x), max(LDMC_G_dat$x), max(RDMC_G_dat$x), max(RDiam_G_dat$x), max(SLA_G_dat$x)),
  y1 = c(1,1,1,1,1)
)


#make a multipanel figure that shows only the graminoid survival probs for 3 traits
gramSurvFigure <- ggplot(data = GramDat) +
  geom_ribbon(aes(x = x, ymin = CI_low, ymax = CI_high, fill = SPEI), alpha = 0.3) +
  geom_line(aes(x=x, GramSurv, col = SPEI))  + 
  geom_rug(aes(x = rug), data = RugDat_G) +
  labs(title = "Graminoid Survival") +
  xlab(NULL) +
  ylab("Probability of Graminoid Survival") +
  scale_y_continuous(limits = c(0,1)) +
  scale_color_manual(labels = c("dry year", "wet year"), values = c("#fc8d62","#66c2a5")) +
  scale_fill_manual(values = c("#fc8d62","#66c2a5"), guide = FALSE) +
  facet_wrap(.~factor(trait, levels = c("scaled(Turgor Loss Point) (MPa)", "scaled(Leaf Dry Matter Content) (g/g)", "scaled(Root Dry Matter Content) (g/g)","scaled(Average Root Diameter) (cm)", "scaled(Specific Leaf Area) (g/cm2)")), scales = "free_x", strip.position =  "bottom", ncol =1) +
  theme_classic()+
  theme(legend.position = "bottom", legend.title = element_blank(), legend.background = element_rect(fill="grey95",size=0.5, linetype="solid"), strip.background = element_rect(colour=NA, fill=NA), strip.placement = "outside", strip.text.x = element_text(margin = margin(0, 0, 1.5, 0)), plot.title = element_text(hjust = 0.5, size = 13, face = "bold")) +
  geom_text(data= dat_text, mapping = aes(x = x, y = y, label = label), size = 3, fontface = "bold") +
  geom_text(data= dat_text, mapping = aes(x = x1, y = y1, label = sig), size = 3.5, fontface = "bold")

##make figure for forb survival
#get 2.5 and 97.5 percentiles of the distribution
meanSPEI_F <- mean(CO_point_all$SPEI_s)
sdSPEI_F <- sd(CO_point_all$SPEI_s)
#get 97.5 quantile of the distribution
SPEI_97_5_F <- qnorm(.975, meanSPEI_F, sdSPEI_F) 
SPEI_2_5_G <- qnorm(.025, meanSPEI_F, sdSPEI_F)

spei_vals <- c(SPEI_2_5_G, SPEI_97_5_G)

#for TLP_s
TLP_vals <- seq(min(CO_point_all$TLP_s, na.rm = TRUE), max(CO_point_all$TLP_s, na.rm = TRUE), length.out = 20)
TLP_F_dat <- ggpredict(mSurvTLP_forbs, terms = c("TLP_s[TLP_vals]", "SPEI_s[spei_vals]"), type = "fixed", back.transform = TRUE)


#for LDMC_s
LDMC_vals <- seq(min(CO_point_all$LDMC_s, na.rm = TRUE), max(CO_point_all$LDMC_s, na.rm = TRUE), length.out = 20)
LDMC_F_dat <- ggpredict(mSurvLDMC_forbs, terms = c("LDMC_s[LDMC_vals]", "SPEI_s[spei_vals]"), type = "fixed", back.transform = TRUE)

#for RDMC_s
RDMC_vals <- seq(min(CO_point_all$RDMC_s, na.rm = TRUE), max(CO_point_all$RDMC_s, na.rm = TRUE), length.out = 20)
RDMC_F_dat <- ggpredict(mSurvRDMC_forbs, terms = c("RDMC_s[RDMC_vals]", "SPEI_s[spei_vals]"), type = "fixed", back.transform = TRUE)

#for RDiam_s
RDiam_vals <- seq(min(CO_point_all$RDiam_s, na.rm = TRUE), max(CO_point_all$RDiam_s, na.rm = TRUE), length.out = 20)
RDiam_F_dat <- ggpredict(mSurvRDiam_forbs, terms = c("RDiam_s[RDiam_vals]", "SPEI_s[spei_vals]"), type = "fixed", back.transform = TRUE)

#for SLA
SLA_vals <- seq(min(CO_point_all$SLA_s, na.rm = TRUE), max(CO_point_all$SLA_s, na.rm = TRUE), length.out = 20)
SLA_F_dat <- ggpredict(mSurvSLA_forbs, terms = c("SLA_s[SLA_vals]", "SPEI_s[spei_vals]"), type = "fixed", back.transform = TRUE)


#make a data.frame to contain all of the values for each trait
ForbDat <- data.frame(trait = c("scaled(Turgor Loss Point) (MPa)"), x = TLP_F_dat$x, ForbSurv = TLP_F_dat$predicted, CI_low = TLP_F_dat$conf.low, CI_high = TLP_F_dat$conf.high, SPEI = TLP_F_dat$group, lab = "G")

ForbDat <- rbind(ForbDat, data.frame(trait = c("scaled(Leaf Dry Matter Content) (g/g)"), x = LDMC_F_dat$x, ForbSurv = LDMC_F_dat$predicted, CI_low = LDMC_F_dat$conf.low, CI_high = LDMC_F_dat$conf.high, SPEI = LDMC_F_dat$group, lab = "H"))

ForbDat <- rbind(ForbDat, data.frame(trait = c("scaled(Root Dry Matter Content) (g/g)"), x = RDMC_F_dat$x, ForbSurv = RDMC_F_dat$predicted, CI_low = RDMC_F_dat$conf.low, CI_high = RDMC_F_dat$conf.high, SPEI = RDMC_F_dat$group, lab = "I"))

ForbDat <- rbind(ForbDat, data.frame(trait = c("scaled(Average Root Diameter) (cm)"), x = RDiam_F_dat$x, ForbSurv = RDiam_F_dat$predicted, CI_low = RDiam_F_dat$conf.low, CI_high = RDiam_F_dat$conf.high, SPEI = RDiam_F_dat$group, lab = "I"))

ForbDat <- rbind(ForbDat, data.frame(trait = c("scaled(Specific Leaf Area) (g/cm2)"), x = SLA_F_dat$x, ForbSurv = SLA_F_dat$predicted, CI_low = SLA_F_dat$conf.low, CI_high = SLA_F_dat$conf.high, SPEI = SLA_F_dat$group, lab = "I"))

#make data for rug plot
RugDat_F <-  data.frame(rug = CO_point_all$TLP_s, trait = "scaled(Turgor Loss Point) (MPa)")
RugDat_F <- rbind(RugDat_F, data.frame(rug = CO_point_all$LDMC_s, trait = "scaled(Leaf Dry Matter Content) (g/g)"))
RugDat_F <- rbind(RugDat_F, data.frame(rug = CO_point_all$RDMC_s, trait = "scaled(Root Dry Matter Content) (g/g)"))
RugDat_F <- rbind(RugDat_F, data.frame(rug = CO_point_all$RDiam_s, trait = "scaled(Average Root Diameter) (cm)"))
RugDat_F <- rbind(RugDat_F, data.frame(rug = CO_point_all$SLA_s, trait = "scaled(Specific Leaf Area) (g/cm2)"))

#text for labels
dat_text <- data.frame(
  label = c("C", "F", "I", "L", "O"),
  sig = c("*", "*,§","*,§", "", "*,§" ),
  trait = c("scaled(Turgor Loss Point) (MPa)", "scaled(Leaf Dry Matter Content) (g/g)", "scaled(Root Dry Matter Content) (g/g)", "scaled(Average Root Diameter) (cm)", "scaled(Specific Leaf Area) (g/cm2)"),
  x    = c(min(TLP_F_dat$x),min(LDMC_F_dat$x),min(RDMC_F_dat$x), min(RDiam_F_dat$x), min(SLA_F_dat$x)),
  y     = c(1,1,1, 1, 1),
  x1 = c(max(TLP_F_dat$x), max(LDMC_F_dat$x), max(RDMC_F_dat$x), max(RDiam_F_dat$x), max(SLA_F_dat$x)),
  y1 = c(1,1,1,1,1)
)


#make a multipanel figure that shows only the graminoid survival probs for 5 traits
forbSurvFigure <- ggplot(data = ForbDat) +
  geom_ribbon(aes(x = x, ymin = CI_low, ymax = CI_high, fill = SPEI), alpha = 0.3) +
  geom_line(aes(x=x, ForbSurv, col = SPEI))  + 
  geom_rug(aes(x = rug), data = RugDat_F) +
  labs(title = "Forb Survival") +
  xlab(NULL) +
  ylab("Probability of Forb Survival") +
  scale_y_continuous(limits = c(0,1)) +
  scale_color_manual(labels = c("dry year", "wet year"), values = c("#fc8d62","#66c2a5")) +
  geom_text(data= dat_text, mapping = aes(x = x, y = y, label = label), size = 3, fontface = "bold") +
  geom_text(data= dat_text, mapping = aes(x = x1, y = y1, label = sig), size = 3.5, fontface = "bold") +
  scale_fill_manual(values = c("#fc8d62","#66c2a5"), guide = FALSE) +
  facet_wrap(~factor(trait, levels = c("scaled(Turgor Loss Point) (MPa)", "scaled(Leaf Dry Matter Content) (g/g)", "scaled(Root Dry Matter Content) (g/g)","scaled(Average Root Diameter) (cm)", "scaled(Specific Leaf Area) (g/cm2)")), scales = "free_x", strip.position =  "bottom", ncol = 1) +
  theme_classic()+
  theme(legend.position = "bottom", legend.title = element_blank(), legend.background = element_rect(fill="grey95",size=0.5, linetype="solid"), strip.background = element_rect(colour=NA, fill=NA), strip.placement = "outside", strip.text.x = element_text(margin = margin(0, 0, 1.5, 0)), plot.title = element_text(hjust = 0.5, size = 13, face = "bold"))

##make figure for graminoid growth models 
#get 2.5 and 97.5 percentiles of the distribution
meanSPEI <- mean(CO_grow_TLP$SPEI_s)
sdSPEI <- sd(CO_grow_TLP$SPEI_s)
#get 97.5 quantile of the distribution
SPEI_97_5 <- qnorm(.975, meanSPEI, sdSPEI) 
SPEI_2_5 <- qnorm(.025, meanSPEI, sdSPEI)

spei_vals <- c(SPEI_2_5, SPEI_97_5)

#for TLP_s
TLP_vals <- seq(min(CO_grow_TLP$TLP_s, na.rm = TRUE), max(CO_grow_TLP$TLP_s, na.rm = TRUE), length.out = 20)
TLP_grow_dat <- ggpredict(mGrowTLP, terms = c("TLP_s[TLP_vals]", "SPEI_s[spei_vals]"), type = "fixed")

#for LDMC_s
LDMC_vals <- seq(min(CO_grow_LDMC$LDMC_s, na.rm = TRUE), max(CO_grow_LDMC$LDMC_s, na.rm = TRUE), length.out = 20)
LDMC_grow_dat <- ggpredict(mGrowLDMC, terms = c("LDMC_s[LDMC_vals]", "SPEI_s[spei_vals]"), type = "fixed")

#for RDMC_s
RDMC_vals <- seq(min(CO_grow_RDMC$RDMC_s, na.rm = TRUE), max(CO_grow_RDMC$RDMC_s, na.rm = TRUE), length.out = 20)
RDMC_grow_dat <- ggpredict(mGrowRDMC, terms = c("RDMC_s[RDMC_vals]", "SPEI_s[spei_vals]"), type = "fixed")

#for RDiam_s
RDiam_vals <- seq(min(CO_grow_RDiam$RDiam_s, na.rm = TRUE), max(CO_grow_RDiam$RDiam_s, na.rm = TRUE), length.out = 20)
RDiam_grow_dat <- ggpredict(mGrowRDiam, terms = c("RDiam_s[RDiam_vals]", "SPEI_s[spei_vals]"), type = "fixed")

#for SLA_s
SLA_vals <- seq(min(CO_grow_SLA$SLA_s, na.rm = TRUE), max(CO_grow_SLA$SLA_s, na.rm = TRUE), length.out = 20)
SLA_grow_dat <- ggpredict(mGrowSLA, terms = c("SLA_s[SLA_vals]", "SPEI_s[spei_vals]"), type = "fixed")


#make a data.frame to contain all of the values for each trait
GrowthDat <- data.frame(trait = c("scaled(Turgor Loss Point) (MPa)"), x = TLP_grow_dat$x, Growth = TLP_grow_dat$predicted, CI_low = TLP_grow_dat$conf.low, CI_high = TLP_grow_dat$conf.high, SPEI = TLP_grow_dat$group)
GrowthDat <- rbind(GrowthDat, data.frame(trait = c("scaled(Leaf Dry Matter Content) (g/g)"), x = LDMC_grow_dat$x, Growth = LDMC_grow_dat$predicted, CI_low = LDMC_grow_dat$conf.low, CI_high = LDMC_grow_dat$conf.high, SPEI = LDMC_grow_dat$group))
GrowthDat <- rbind(GrowthDat, data.frame(trait = c("scaled(Root Dry Matter Content) (g/g)" ), x = RDMC_grow_dat$x, Growth = RDMC_grow_dat$predicted, CI_low = RDMC_grow_dat$conf.low, CI_high = RDMC_grow_dat$conf.high, SPEI = RDMC_grow_dat$group))
GrowthDat <- rbind(GrowthDat, data.frame(trait = c("scaled(Average Root Diameter) (cm)"), x = RDiam_grow_dat$x, Growth = RDiam_grow_dat$predicted, CI_low = RDiam_grow_dat$conf.low, CI_high = RDiam_grow_dat$conf.high, SPEI = RDiam_grow_dat$group))
GrowthDat <- rbind(GrowthDat, data.frame(trait = c("scaled(Specific Leaf Area) (g/cm2)"), x = SLA_grow_dat$x, Growth = SLA_grow_dat$predicted, CI_low = SLA_grow_dat$conf.low, CI_high = SLA_grow_dat$conf.high, SPEI = SLA_grow_dat$group))

#make a data.frame with data for the rug plot
RugDat <- data.frame(rug = CO_grow_TLP$TLP_s, trait = "scaled(Turgor Loss Point) (MPa)")
RugDat <- rbind(RugDat, data.frame(rug = CO_grow_LDMC$LDMC_s, trait = "scaled(Leaf Dry Matter Content) (g/g)"))
RugDat <- rbind(RugDat, data.frame(rug = CO_grow_RDMC$RDMC_s, trait = "scaled(Root Dry Matter Content) (g/g)"))
RugDat <- rbind(RugDat, data.frame(rug = CO_grow_RDiam$RDiam_s, trait = "scaled(Average Root Diameter) (cm)"))
RugDat <- rbind(RugDat, data.frame(rug = CO_grow_SLA$SLA_s, trait = "scaled(Specific Leaf Area) (g/cm2)"))


#text for labels
dat_text <- data.frame(
  label = c("B", "E", "H", "K","N"),
  sig = c("", "","", "", "" ), 
  trait = c("scaled(Turgor Loss Point) (MPa)", "scaled(Leaf Dry Matter Content) (g/g)", "scaled(Root Dry Matter Content) (g/g)", "scaled(Average Root Diameter) (cm)", "scaled(Specific Leaf Area) (g/cm2)"),
  x    = c(min(TLP_grow_dat$x), min(LDMC_grow_dat$x), min(RDMC_grow_dat$x), min(RDiam_grow_dat$x), min(SLA_grow_dat$x)),
  y     = c(4.4,4.4,4.4,4.4,4.4),
  x1 = c(max(TLP_grow_dat$x), max(LDMC_grow_dat$x), max(RDMC_grow_dat$x), max(RDiam_grow_dat$x), max(SLA_grow_dat$x)),
  y1 = c(4.4,4.4,4.4,4.4,4.4)
)

#make a multipanel figure
GrowthExtraFig <- ggplot(data = GrowthDat) +
  geom_ribbon(aes(x = x, ymin = CI_low, ymax = CI_high, fill = SPEI), alpha = 0.3) +
  geom_line(aes(x=x, Growth, col = SPEI))  + 
  geom_rug(aes(x = rug), data = RugDat) +
  labs(title = "Graminoid Growth") +
  xlab(NULL) +
  ylab(expression(size[t+1])) +
  scale_color_manual(labels = c("dry year", "wet year"), values = c("#fc8d62","#66c2a5")) +
  scale_fill_manual(values = c("#fc8d62","#66c2a5"), guide = FALSE) +
  geom_text(data= dat_text, mapping = aes(x = x, y = y, label = label), size = 3, fontface = "bold") +
  geom_text(data = dat_text, mapping = aes(x = x1, y = y1, label = sig), size = 3.5, fontface = "bold") +
  facet_wrap(.~factor(trait, levels = c("scaled(Turgor Loss Point) (MPa)", "scaled(Leaf Dry Matter Content) (g/g)", "scaled(Root Dry Matter Content) (g/g)","scaled(Average Root Diameter) (cm)", "scaled(Specific Leaf Area) (g/cm2)")), scales = "free_x", strip.position =  "bottom", ncol = 1) +
  theme_classic()+
  theme(legend.position = "bottom", legend.title = element_blank(), legend.background = element_rect(fill="grey95",size=0.5, linetype="solid"), strip.background = element_rect(colour=NA, fill=NA), strip.placement = "outside", strip.text.x = element_text(margin = margin(0, 0, 1.5, 0)), plot.title = element_text(hjust = 0.5, size = 13, face = "bold"))

mainObs <- cowplot::plot_grid( gramSurvFigure, GrowthExtraFig, forbSurvFigure, ncol = 3, align = "h", axis = "tb", rel_widths = c(.9,1, .9)) %>% 
  cowplot::ggdraw() + 
  cowplot::draw_grob(grid::rectGrob(gp = grid::gpar(fill = "white", col = "white")) ,hjust = 0, vjust = 0, x = .07, y = .059, width = .92, height = .0227) +
  cowplot::draw_grob(grid::rectGrob(gp = grid::gpar(fill = "white", col = "white")) ,hjust = 0, vjust = 0, x = .07, y = .006, width = .33, height = .05) + #cover up left legend 
  cowplot::draw_grob(grid::rectGrob(gp = grid::gpar(fill = "white", col = "white")) ,hjust = 0, vjust = 0, x = .75, y = .006, width = .33, height = .05) #cover up right legend

mainObs <- ggdraw(add_sub(mainObs, label = bquote("scaled(Specific Leaf Area) (c"*m^2*"/g)"),x = .5, y = 3, size = 9))
mainObs <- ggdraw(add_sub(mainObs, label = bquote("scaled(Specific Leaf Area) (c"*m^2*"/g)"),x = .19, y = 4, size = 9))
mainObs <- ggdraw(add_sub(mainObs, label = bquote("scaled(Specific Leaf Area) (c"*m^2*"/g)"),x = .873, y = 5, size = 9))
mainObs
```

## Here is the model output for these new growth models: 
```{r, echo = FALSE}
stargazer(mGrowTLP, mGrowLDMC, mGrowSLA, mGrowRDMC, mGrowRTD, mGrowSRL, mGrowRDiam, type = "text", style = "all2", column.labels = c("TLP", "LDMC", "SLA", "RDMC","RTD", "SRL", "RDiam"), dep.var.labels = c("log(size t+1)"), digits = 2, model.numbers = FALSE, report = c("vc*"), omit = c("TLP_s", "LDMC_s", "SLA_s", "RDMC_s", "RTD_s", "SRL_s", "RDiam_s"), 
          add.lines =  list(
            Trait = c("Trait", "-0.17**", "0.13", "0.05", "0.05", "-0.20**", "-0.09", "0.02") ,
            Blank = c("", "", "", "", "", "", "", ""),
            TraitBySPEI = c("Trait:SPEI", "0.01","-0.02", "-0.01", "-0.01", "0.01", "0.01", "-0.016"),
            Blank = c("", "", "", "", "", "", "", ""),
            deltaAIC <- c("Delta AIC", "-11.02", "-10.39", "-13.15", "-12.25", "-9.56", "-12.66", "-12.39")
                            ), omit.table.layout = c("-"),
          omit.stat = c("bic", "ll"))
```

## *However* there is a problem with this modeling approach that I'm not sure how to solve. 
If we look at the relationship between trait values and size, we see a negative relationship between TLP and size$_{t+1}$, such that larger plants tend to have a more negative TLP. So this could get in the way of identifying the interactive effect of SPEI and trait on size in year$_{t+1}$. 
```{r}
#fit a model
modSizeTLP <- lm(CO_grams$area_tplus1_s ~ CO_grams$TLP_s + CO_grams$area_s)

ggplot(data = CO_grams) +
  geom_point(aes(x = TLP_s, y = area_tplus1_s), colour = "darkblue", pch = 1) +
  xlab("TLP") +
  ylab("size_t+1") + 
  geom_smooth(aes(x = TLP_s, y = area_tplus1_s),  method = "lm", col = "black", lwd = .75) + 
  theme_classic()
summary(modSizeTLP)
```


# Comparing size_t and size_t+1 for each species
```{r}
## use CO_poly_growth dataset
ggplot(data = CO_grow_TLP) +
  #geom_point(aes(x = area_s, y = area_tplus1_s, col = species)) +
  geom_smooth(aes(x = size_t_log, y = size_tplus1_log, col = species), method = "lm", se = FALSE) +
  theme_classic() +
  xlab(expression(ln(size[italic(t)]))) +
  ylab(expression(ln(size[italic(t+1)]))) +
  theme(legend.text = element_text(face = "italic")) +
  guides(col = guide_legend(title = "Species")) +
  scale_color_brewer(palette = "Set2" )
```

# Comparing size_t and growth for each species
```{r}
## use CO_poly_growth dataset
ggplot(data = CO_grow_TLP) +
  #geom_point(aes(x = size_t_log, y = logDiffArea, col = species)) +
  geom_smooth(aes(x = size_t_log, y = logDiffArea, col = species), method = "gam", se = FALSE) +
  theme_classic() +
  xlab("size_t") +
  ylab("Log Difference of Area (t to t+1)")
```

# Interaction between SPEI and size t+1 
### color coded by trait value
```{r}
plot(Effect(focal.predictors=c("SPEI_s", "species"), mod = mGrowTLP))

sjPlot::sjp.lmer(mGrowTLP, vars = "SPEI_s", type = "ri.slope")
```

Compare SPEI and size_t+1 accross a spectrum of TLP for graminoids 
```{r}
#get unique values of TLP 
trait_vals <- sort(unique(CO_grow_TLP$TLP_s))

#for TLP_s
SPEI_vals <- seq(min(CO_grow_TLP$SPEI_s, na.rm = TRUE), max(CO_grow_TLP$SPEI_s, na.rm = TRUE), length.out = 20)
TLP_grow_dat <- ggpredict(mGrowTLP, terms = c( "SPEI_s[SPEI_vals]", "TLP_s[trait_vals]"), type = "fixed")

#make a data.frame to contain all of the values for each trait
SPEIDat <- data.frame(trait = c("scaled(Turgor Loss Point) (MPa)"), x = TLP_grow_dat$x, Growth = TLP_grow_dat$predicted, CI_low = TLP_grow_dat$conf.low, CI_high = TLP_grow_dat$conf.high, TLP = TLP_grow_dat$group)

#make a data.frame with data for the rug plot
RugDat <- data.frame(rug = CO_grow_TLP$SPEI_s, trait = "scaled(SPEI)")

#make a figure
 ggplot(data = SPEIDat) +
  geom_ribbon(aes(x = x, ymin = CI_low, ymax = CI_high, fill = TLP), alpha = 0.3) +
  geom_line(aes(x=x, Growth, col = TLP))  + 
  geom_rug(aes(x = rug), data = RugDat) +
  labs(title = "Graminoid Growth") +
  xlab("SPEI") +
  ylab(expression(size[t+1])) +
  scale_color_manual(labels = c("Buchloe dactyloides", "Schedonnardus paniculatus", "Bouteloua gracilis", "Stipa comata", "Aristida longiseta" , "Sporobolus cryptandrus", "Sitanion hystrix"), values = scales::hue_pal(h = c(0, 180))(7)) +
  scale_fill_manual(values = scales::hue_pal(h = c(0, 180))(7)) +
  theme_classic()
```
Compare SPEI and P(survival) accross a spectrum of TLP for graminoids 
```{r}
#get unique values of TLP 
trait_vals <- sort(unique(CO_poly_TLP$TLP_s))

#for TLP_s
SPEI_vals <- seq(min(CO_poly_TLP$SPEI_s, na.rm = TRUE), max(CO_poly_TLP$SPEI_s, na.rm = TRUE), length.out = 20)
TLP_grow_dat <- ggpredict(mSurvTLP_grams, terms = c( "SPEI_s[SPEI_vals]", "TLP_s[trait_vals]"), type = "fixed")

#make a data.frame to contain all of the values for each trait
SPEIDat <- data.frame(trait = c("scaled(Turgor Loss Point) (MPa)"), x = TLP_grow_dat$x, Growth = TLP_grow_dat$predicted, CI_low = TLP_grow_dat$conf.low, CI_high = TLP_grow_dat$conf.high, TLP = TLP_grow_dat$group)

#make a data.frame with data for the rug plot
RugDat <- data.frame(rug = CO_poly_TLP$SPEI_s, trait = "scaled(SPEI)")

#make a figure
 ggplot(data = SPEIDat) +
  geom_ribbon(aes(x = x, ymin = CI_low, ymax = CI_high, fill = TLP), alpha = 0.1) +
  geom_line(aes(x=x, Growth, col = TLP))  + 
  geom_rug(aes(x = rug), data = RugDat) +
  labs(title = "") +
  xlab("SPEI") +
  ylab("P(graminoid survival)") +
  ylim(c(0,1)) +
  scale_color_manual(labels = c("Buchloe dactyloides", "Schedonnardus paniculatus", "Carex eleocharis", "Bouteloua gracilis", "Stipa comata", "Aristida longiseta" , "Sporobolus cryptandrus", "Sitanion hystrix"), values = scales::hue_pal(h = c(0, 180))(8)) +
  scale_fill_manual(values = scales::hue_pal(h = c(0, 180))(8)) +
  theme_classic()
```

# Comparing Unique vs. Uniform SPEI  (for graminoids only)s
The following results are for graminoids only

```{r, echo = FALSE}
temp <- CO_grams[,c("species", "year_t", "SPEI_unique", "SPEI_uniform")]
gramsSPEI_unique <- aggregate(temp[,c("SPEI_unique", "SPEI_uniform")], by = list(temp$species, temp$year_t), FUN = mean) %>% 
  rename(species = Group.1, year_t = Group.2) %>% 
  arrange(species) %>% 
  filter(is.na(SPEI_unique) == FALSE) %>% 
  mutate(year_num = as.numeric(as.character(year_t)))

## get photosynthetic pathway data
CO_traits_temp <- CO_traits[,c("species", "C3_C4")]
## merge w/ SPEI data
gramsSPEI_unique <- left_join(gramsSPEI_unique, CO_traits_temp)
#add additional pathway data
gramsSPEI_unique[gramsSPEI_unique$species=="Schedonnardus paniculatus", "C3_C4"] <- "C4"
gramsSPEI_unique[gramsSPEI_unique$species=="Muhlenbergia torreyi", "C3_C4"] <- "C4"
gramsSPEI_unique[gramsSPEI_unique$species=="Carex spp.", "C3_C4"] <- "C3"
gramsSPEI_unique[gramsSPEI_unique$species=="Carex filifolia", "C3_C4"] <- "C3"

gramsSPEI_unique <- gramsSPEI_unique %>% 
  mutate(C3_C4 = as.factor(C3_C4))
```

## Below are the *unique* SPEI values for each species

The black line shows SPEI when calculated for a uniform interval each year

```{r, echo = FALSE}
## plot the unique SPEI values (grams only)
# plotted by species
ggplot(data = gramsSPEI_unique) +
  geom_line(aes(x = year_num, y = SPEI_unique, col = C3_C4, lty = species)) +
  geom_line(aes(x = year_num, y = SPEI_uniform)) +
  theme_classic()
```

## Below are the *unique* SPEI values for each species, color-coded by photosynthetic pathway (graminoids only)
The black asterisks indicate the value for SPEI when calculated for a uniform interval each year. 
These patterns are somewhat surprising, since in many years C4 species actually tend to have a more positive SPEI (wetter) than C3 species, which is opposite of what we would expect. 
```{r, echo = FALSE}
# plotted by photosynthetic type (grams only)
ggplot(data = gramsSPEI_unique) +
  geom_boxplot(aes(x = year_t, y = SPEI_unique, col = C3_C4)) +
geom_point(aes(x = year_t, y = SPEI_uniform), col= "black", shape = 8) +
  theme_classic()
```

## For species-level unique SPEI, year is a significant predictor but species is not. 
```{r}
## is the variation between years greater than the variation between species in one year?
testM <- lm(SPEI_unique ~ year_t + species, data = gramsSPEI_unique)
anova(testM)
#year is a more important predictor than species (is significant, while species isn't)
```

## When photosynthetic pathway is considered as well, year is still the most significant predictor of SPEI. Photosynthetic pathway has a slightly significant effect, while species has no significant effect. 
```{r}
testM_photo <- lm(SPEI_unique ~ year_t + C3_C4 + species, data = gramsSPEI_unique)
anova(testM_photo)
```
### These results increase my confidence in using species-level SPEI, but I'd like to know what you all think.


## These are the yearly precipitation and temperature patterns accross all years included in the demographic dataset
```{r, echo = FALSE}
#read in montly climate data
setwd("/Users/Alice/Dropbox/Grad School/Research/Trait Project/Data/Climate Data/CO Climate/")
CO_clim <- read.csv("./CO_Climate_All.csv", stringsAsFactors = FALSE)
#format date column correctly
CO_clim$Date <- as.POSIXct(CO_clim$Date, tz = "UTC", format = "%m/%d/%y")
#get only the years present in the demographic dataset
CO_clim <- CO_clim[lubridate::year(CO_clim$Date) %in% c(2000:2009),]
CO_clim$xInt_1 <- 5
CO_clim$xInt_2 <- 9
#plot precip over months
ggplot(data = CO_clim) + 
  geom_vline(aes(xintercept = xInt_1), col = "lightgrey") +
  geom_vline(aes(xintercept = xInt_2), col = "lightgrey") +
  geom_col(aes(x = lubridate::month(Date), y = Monthly.ppt), fill = "darkblue", col = "darkblue", width = .5) +
  geom_line(aes(x = lubridate::month(Date), y = (Mean.Monthly.T/.3)), col = "red") +
  theme_classic() +
  scale_y_continuous("Precip (mm)", 
                     sec.axis = sec_axis(~ . * .3, name = "Temp (C)"))  + 
  theme(axis.title.y = element_text(color = "darkblue"),
        axis.title.y.right = element_text(color = "red"))  +
  facet_wrap(~factor(lubridate::year(CO_clim$Date))) +
  xlab("Month") 
```
I plotted these patterns to check if it makes sense that sometimes SPEI for C4 species is actually higher (wetter) than for C3 species. It does look like in some years, there is more precip later in the summer. So  maybe that pattern of SPEI does make sense. 



